Load libraries
```{r}
source("../General/general_functions.R") 
library(cowplot)
library(tidyverse)
```
Load data and format
```{r}
map <- read.csv("../Data/210201_natres_meta.csv", header = T, skip = 1)

map <- map %>% 
  rename("SampleID" = "sampleID")

otu <- read.csv("../Data/210201_natres_norm.csv", header = T)

row.names(otu) <- otu$contig
otu <- otu[,-1]
otu <- otu[,colnames(otu) %in% map$SampleID]
```

Calculate spatial distance
```{r}
#Generate a matrix with the coordinates recorded for each sample
mtx <- map %>% 
  select(SampleID,GPS) %>% 
  mutate(GPS = str_replace_all(GPS, "N | W |\\°|\\'", "")) %>% 
  separate(GPS, c("lat", "lon"), sep = ";") %>% 
  mutate(lon = paste("-", lon, sep = "")) %>% 
  mutate(lat_d = as.numeric(measurements::conv_unit(lat, from = 'deg_dec_min', to = 'dec_deg')),
         lon_d = as.numeric(measurements::conv_unit(lon, from = 'deg_dec_min', to = 'dec_deg'))) %>% 
  select(SampleID, lon_d, lat_d) %>% 
  as.data.frame()

row.names(mtx) <- mtx$SampleID
mtx <- mtx[,-1]
mtx <- as.matrix(mtx)
mtx

#Calculate a matrix with the pairwise spatial distance
spatial.dist <- geosphere::distm(mtx, mtx) %>% as.data.frame()
row.names(spatial.dist) <- row.names(mtx)
colnames(spatial.dist) <- row.names(mtx)

#Remove redundant values
spatial.dist[upper.tri(spatial.dist, diag = F)] <- NA 

#Generate a long data frame with pairwise spatial distances
spatial.dist.tidy <- spatial.dist %>% 
  mutate(SampleID.x = row.names(.)) %>% 
  gather(key = "SampleID.y", value = "SpatialDistance", -SampleID.x) %>% 
  mutate(SpatialDistance = SpatialDistance/1000) %>% 
  filter(!is.na(SpatialDistance))

spatial.dist.tidy
```

Calculate Jaccard similarities between samples. There is so little overlap in the vOTUs detected between sites that a presence/absence metric makes more sense than Bray-Curtis
```{r}
#Calculate Jaccard distances
jac <- vegan::vegdist(t(otu), method = "jaccard")
#Convert it to similarities
jac <- 1- jac

#Generate a long data frame with the results
jac.tidy <- jac %>% 
  as.matrix() %>% 
  as.data.frame() %>% 
  mutate(SampleID.x = row.names(.)) %>% 
  gather(key = "SampleID.y", value = "JacSim", -SampleID.x)

#Merge spatial and Jaccard data frames and reformat
spatial.jaccard.tidy <- spatial.dist.tidy %>% 
  inner_join(jac.tidy, by = c("SampleID.x", "SampleID.y")) %>% 
  filter(SpatialDistance > 0) %>% 
  inner_join(map, by = c("SampleID.x" = "SampleID")) %>% 
  inner_join(map, by = c("SampleID.y" = "SampleID")) %>% 
  mutate(HabitatDetection = ifelse(gen_habitat.x == gen_habitat.y, "single habitat", "mutliple habitats"),
         SiteDetection = ifelse(natres_site.x == natres_site.y, "single site", "multiple sites")) %>% 
  mutate(Detection = paste(HabitatDetection, SiteDetection, sep = " / "))
```

Plot figures with spatial distance boxplot and Jaccard rank dot plot. There's two versions: one in which the boxplot only includes pairs of samples with Jaccard similarities > 0, another one with all the data points
```{r}
#Make boxplot showing the distributions of spatial distances excluding all the pairwise comparisons with no overlap in the vOTUs detected (Jaccard Sim > 0)
box.filt.p <- spatial.jaccard.tidy %>% 
  filter(JacSim > 0) %>% 
  ggplot(aes(1, SpatialDistance)) +
  geom_boxplot() + 
  geom_jitter(aes(color = Detection), width = 0.1, height = 0, alpha = 0.8) +
  scale_color_manual(values = c("dodgerblue4", "dodgerblue1", "goldenrod3", "goldenrod1")) +
  xlab("") +
  ylab("Pairwise spatial distance (km)") +
  ylim(-10,120) +
  theme_minimal() +
  theme(text = element_text(size = 15),
        axis.text.x = element_blank(),
        legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())

#Make boxplot showing the distributions of spatial distances with all the data points
box.all.p <- spatial.jaccard.tidy %>% 
  ggplot(aes(1, SpatialDistance)) +
  geom_boxplot() + 
  geom_jitter(aes(color = Detection), width = 0.1, height = 0, alpha = 0.8) +
  scale_color_manual(values = c("dodgerblue4", "dodgerblue1", "goldenrod3", "goldenrod1")) +
  xlab("") +
  ylab("Pairwise spatial distance (km)") +
  ylim(-10,120) +
  theme_minimal() +
  theme(text = element_text(size = 15),
        axis.text.x = element_blank(),
        legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())

#Make a plot showing the rank of each pairwise comparison based on the similarity between communities, color-coded by whether the comparison involved samples between multiple/single habitats/sites
dot.rank.p <- spatial.jaccard.tidy %>% 
  filter(JacSim > 0) %>%
  mutate(Rank = rank(-JacSim)) %>%
  ggplot(aes(Rank, SpatialDistance)) +
  geom_point(aes(size = JacSim, color = Detection), shape = 16, alpha = 0.8) +
  scale_color_manual(name = "Group", values = c("dodgerblue4", "dodgerblue1", "goldenrod3", "goldenrod1")) +
  xlab("Community similarity rank") +
  ylim(-10,120) +
  labs(size = "Jaccard similarity") +
  theme_minimal() +
  theme(text = element_text(size = 15),
        axis.text.y = element_blank(),
        axis.title.y = element_blank())

plot_grid(box.all.p, dot.rank.p, nrow = 1, rel_widths = c(1,4), align = "vh", axis = "bt")
plot_grid(box.filt.p, dot.rank.p, nrow = 1, rel_widths = c(1,4), align = "vh", axis = "bt")
```

Plot figures with spatial distance boxplot, Jaccard similarity dot plot, and Jaccard similarity histogram. There's two versions: one in which the boxplot only includes pairs of samples with Jaccard similarities > 0, another one with all the data points
```{r}
#Make a plot showing the Jaccard similarity (log transformed) between pairs of communities, color-coded by whether the comparison involved samples between multiple/single habitats/sites
dot.jac.p <- spatial.jaccard.tidy %>% 
  filter(JacSim > 0) %>%
  ggplot(aes(-log10(JacSim), SpatialDistance)) +
  geom_point(aes(color = Detection), shape = 16, alpha = 0.8) +
  scale_color_manual(name = "Group", values = c("dodgerblue4", "dodgerblue1", "goldenrod3", "goldenrod1")) +
  xlab("Jaccard similarity\n(-log10)") +
  ylim(-10,120) +
  theme_minimal() +
  theme(text = element_text(size = 15),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none")

#Make a histogram showing the distribution of types of samples across the spatial distance data. Include only comparisons where Jaccard similarity > 0
hist.p <- spatial.jaccard.tidy %>% 
  filter(JacSim > 0) %>%
  ggplot(aes(SpatialDistance)) +
  geom_histogram(aes(fill = Detection), binwidth = 12) +
  scale_fill_manual(name = "Group", values = c("dodgerblue4", "dodgerblue1", "goldenrod3", "goldenrod1")) +
  ylab("Number of\npaired samples") +
  xlim(-10,120) +
  labs(size = "Jaccard similarity") +
  theme_minimal() +
  theme(text = element_text(size = 15),
        axis.text.y = element_blank(),
        axis.title.y = element_blank()) +
  coord_flip()

plot_grid(box.all.p, dot.jac.p, hist.p, nrow = 1, align = "vh", rel_widths = c(1,2,4))
plot_grid(box.filt.p, dot.jac.p, hist.p, nrow = 1, align = "vh", rel_widths = c(1,2,4))
```




